### 2019-06-03 Documents/literature_review/pages/review_content.tex
Changes; - Stated the re-write of the first section.  The previous state has been backed up for reference and persistance

\section{Review - Fingerprint representations}

\subsection{Fingerprint representation comparison}

\textbf{TODO: } - Create written introduction setting the scene.
\textbf{TODO: } - Summarise research questions after lit' review. Back them up using my discussion.
% \textbf{TODO: } - Check evaluation with each paper in regards to:
% \begin{itemize}
%     \item Participant size [P]
%     \item Social aspects, where have they come from? Age? Gender? Occupation?  [S]
%     \item Target entropy [E]
%     \item Attacker strength [A]
% \end{itemize}
% [P, S, E, A]

Current research into human-based validation of fingerprints has almost exclusively focused on the usability of such schemes. The following section will individually discuss available research findings alongside an overall comparison of the research recommendations.

% [A]
Some of the first work in this area was performed by \textbf{Hsu-Chun Hsiao, \textit{et al.}}\cite{hsiao2009study}. The main aim of the study was to provide the first insight into the best encoding schemes used to represent a hash fingerprint. The schemes used were Base32, English words, Random Art\cite{perrig1999hash}, Flag\cite{ellison2003public}, T-Flag\cite{lin2010spate}, Flag Ext\footnote{The authors' own improvement on Flag} and finally Chinese, Korean and Japanese symbolic encodings. A total of 436 participants were assessed in the study.

In addition to the central aims of the paper, the authors intended to provide empirical links between participant attributes (such as age or gender) and performance with the encoding scheme. This is something rarely seen in the related research. Hsu-Chun Hsiao \textit{et al.} also provided two categories of similarities; ``hard'' and ``easy''. ``Hard'' pair only have a very small difference, therefore, difficult to distinguish. These were designed to be the worst and best case respectively. The reasoning behind inclusion of Chinese, Korean and Japanese characters was to investigate the links between language ability and distinguishing subtle differences between encodings in the respective language. The entropy of encodings ranged from 22-bits to 28-bits. This is the lowest level of entropy used in the assessed literature. To quantify this a hash rate of 10 MH/s SHA-1 compressions would be able to find a match for 28-bits in 26.8 seconds. This is addressed by the author at a stage in the paper where they state \textit{``However, increasing entropy is not a solution because it sacrifices usability and accuracy. With more entropy, representations will contain longer sequences of characters or more minute details which will lead to increased time and errors during comparisons.''}. This claim is entirely unsubstantiated with any empirical evidence alongside no discussion into why the level of entropy was chosen. This, however, was followed by in interesting recommendation into the use of commitments\cite{blum1982coin}. This provides an attacker with a single chance to find a collision, thus, providing a valid use case for low levels of entropy.\\
Another limitation is the authors' decision to exclude hexadecimal and numerical encodings (some of the most widespread encoding schemes). This was based on their own claims on similarities to Base32 and ``well-known deficiencies" in the excluded schemes. This point was provided with no further justification or quantified in any way. It is also not consistent with available research, for example, in \textit{``Empirical Study of Textual Key-Fingerprint Representations''}\cite{dechand2016empirical} it was shown numerical representations performed significantly better than that of Base32.

Findings from the paper showed that age and gender do not affect the accuracy of the scheme, however, younger participants were significantly faster. \\
The paper recommends Base32, Random Art, T-Flag or their own improvement; Flag Ext. These encoding recommendations were also supplemented with a review of the requirements required by these schemes. The recommendation of graphical encoding schemes is inconsistent with alternative literature, where most other papers have found graphical encodings easy to use but insecure.
In terms of language comprehension; being able to speak the language assists in the ability to discern the differences between hard pairs. Subsequently, knowledge of the language did not assist in differentiating easy pairs as these had high accuracy regardless of the spoken language.

In conclusion, the paper touched upon a wide variety of topics. Moreover, the consideration into the attributes of participants and consideration into how this would affect their performance with the scheme was a substantial addition. This provides a higher level of validity to the results and is something severely lacking from the literature assessed. However, the age of the study limits the applicability to modern day scenarios, and oversights in the development of the study such as the exclusion of hexadecimal encoding and the lack of information gained on the perceived usability of the scheme from users results in an incomplete study. In addition, the paper failed to consider the quantification of attacker strength and, therefore, failed to discuss the feasibility of these attacks.

% [A]
\textbf{Kainda \textit{et al.}}\cite{kainda2009usability} in 2009 investigated fingerprint representations and comparison methods in the context of using humans as the Out-Of-Band (OOB) channel for secure device pairing. This study was comprehensive and assessed a wide variety of concepts. This is one of the few papers in the literature that assessed the way to compare encodings alongside the representations themselves. For comparison methods the paper looked into ``Compare-and-Confirm'', ``Compare-and-Select'', ``Compare-and-Enter'' and ``Barcode'' scanning. This set includes the most common ways of comparison according to available research. Alongside this, the paper reviews Numerical, Alphanumerical, Words, Sentences, Images, Melodies and Sound (Numerical/Alphanumerical). Again, this in comparison to similar literature is highly comprehensive.

In terms of empirical data gathered by the paper, they reviewed the time completed to asses the representation and the number of security and non-security related errors. Furthermore, the authors also collected user opinion on ease-of-use, satisfaction and confidence of the reviewed schemes. The overall schemes were then quantified using a metric known as the Single Usability Metric (SUM)\cite{sauro2005method}. This provides an elementary way to rank and compare the schemes, thus improving the potential validity of the results. 
However, the one apparent downside to the experimental side of the study is the 30 participants enrolled. This is one of the smallest sizes of enrolment in the literature and thus, puts a possible limit on the reliability and applicability of the obtained results. This limitation was, however,  supplemented by a consideration into the demographics of the participants enrolled. The participants were evenly split between male and female alongside an age range of 18 to 75.

Overall, if usability is the only consideration the paper ranked the comparison methods: Compare-and-Confirm, Compare-and-Enter, Compare-and-Select and Barcode. If, however, the consideration of security alongside usability is required the best is: Compare-and-Enter, Barcode, Compare-and-Confirm and Compare-and-Select. Research has been limited in this area of modes of comparison, however, the results from this paper are consistent with the limited alternative literature. In terms of encoding schemes, the paper ranked Numerical, Alphanumerical and Words as the top three encoding schemes (These were all with Compare-and-Confirm). This recommendation took into consideration the usability and security of the scheme and is relatively consistent with other research where numerical and English language based encoding schemes seem to be the most effective overall.

In conclusion, this paper provided a highly comprehensive review of all the aspects of human fingerprint verification, where it reviewed all the elements involved in the process. Limitations of the paper involved the inconsistent and low range of entropy (20-bits to 40-bits)alongside the size of the participants enrolled for the study.

% TODO: Implement checks below:
\textbf{Ersin Uzun \textit{et al.}}\cite{uzun2007usability} exclusively reviewed comparison methods. This is performed in the context of ``Secure device pairing'' and thus some comparison schemes are tested that are not relevant in a fingerprint comparison context, for example  ``Choose-and-Enter'' (A participant chooses a passphrase from an available list and enters it into the other device). Thus, only relevant comparison scheme and their results have been extracted and compared.

The relevant comparison methods tested were, 
``Compare-and-Confirm'', ``Compare-and-Select'' and ``Compare-and-Enter''. These are almost exactly the same encoding schemes assessed in the relevant research literature. The paper also had two rounds of experiments including 40 participants each where changes were made to UX and format between these studies. 
Due to the ``Secure device pairing'' context, consideration needs to be made into the actual strings being encoded. Due to secure device pairing commonly utilising Short Authentication Strings (SAS) the
paper only reviewed encodings of 4 digit long strings. This, therefore, limits the applicability of the results in comparison to other fingerprint encoding literature. Between the two rounds, the authors dropped Confirm-and-Enter due to its low usability and over similarity to a similar scheme called ``Copy'' (used for passphrase verification). The low rating for Confirm-and-Enter is in contrast with research performed by Kainda \textit{et al.}\cite{kainda2009usability} where they ranked Confirm-and-Enter as best method overall. With changes to the UX of Compare-and-Confirm backed up by external research \cite{palmer1990attentional}\cite{hammer2009category} the second round of research was performed with a new set of participants. The alterations took Compare-and-Confirm from 20\% to 0\% fatal error rate (FER). The same trend for Confirm-and-Select with it reducing its FER from 12.5\% to 5\%. The authors, therefore, recommended ``Compare-and-Confirm'' as the overall best choice for comparison methods. Comparing this research to the work done by Kainda \textit{et al.}\cite{kainda2009usability}; this is consistent with their findings where they ranked ``Compare-and-Confirm'' as superior to that of ``Compare-and-Select'' in all regards.

Overall, the paper provides a sizeable insight into suitable methods of comparison, this is due to the cross over between fingerprint comparison and secure device pairing. This paper is limited in terms of the very short strings compared, the different overall research aim and the size of the number of participants used in the studies. In terms of overall recommendations, the results were consistent with that of alternative research with the exception of recommendations regarding ``Compare-and-Enter''. This, however, may be due to the differing use-cases of the comparison methods in the two respective studies.

\textbf{Dechand \textit{et al.}}\cite{dechand2016empirical} empirically investigated the usability of 4 distinct textual representations evaluated with an experiment involving a total of 1047 participants. The textual representations were: Alphanumeric, Numeric, Words and Sentences. They assessed the number of attacks missed with each scheme alongside results from a questionnaire on the participants preferred scheme and perceived usability.

The paper touches upon issues with decentralised methods of identification such a PGP's Web of Trust and the problems these solutions have with user adoption. These points are made in an attempt to validate the requirement for manual comparison of key based fingerprints. This is a common theme that appears in the majority of the reviewed papers. Other references made within the paper touch upon the vulnerabilities and usability issues present with the way humans interact with the security systems. Examples of these are studies showing humans find it difficult to comprehend long and ``meaningless" strings and the lack of actual comparison performed by users in live scenarios. These are, however, acknowledged as limitations in the later stages of the paper. The paper has defined the upper and lower bound costs of the attacker's resources and strength as \$610K to \$16B, with an ability to control 80-bits of the fingerprint. This in comparison to other papers is high and is almost encroaching into the realm of a highly sophisticated attacker. Therefore, the lack of consideration for the lower-end of the attack resource spectrum can be considered another limitation.

Summarised findings from the paper state that conventional encodings such as Hexadecimal and Base32 perform worse than all other alternatives in a realistic threat model with over 10\% of users failing to detect an attack on these encoding schemes. The recommendations of the authors are to replace these encoding schemes with Words or Sentences due to their very high success rate and high usability scores. The performance of the Alphanumerical encoding schemes is relatively consistent with the literature, however, it is hard to say conclusively due to the small number of total schemes assessed making it hard to place alphanumerical encodings in an overall rating.

\textbf{J Tan \textit{et al.}}\cite{tan2017can} work is very similar to that of the research already discussed. Eight distinct fingerprint representations were tested with over 661 participants. Each representation was tested using ``Compare-and-Confirm'' and ``Compare-and-Select''. The small difference with this paper is the inclusion of two novel graphical encodings Vash\footnote{https://github.com/thevash/vash} and Unicorns\footnote{https://unicornify.pictures/}. The compared schemes were: Hexadecimal, Alternating vowel/consonants, Words, Numbers, Sentences, OpenSSH visual host key, Vash and Unicorns.

The target security level chosen was an entropy of 160-bits. This is a very high level of security in comparison to alternative literature. Alongside this, the paper defined attack strengths of an average $2^{60}$ with other specific use-cases for lower and higher attacker strength. Users were recruited via MTurk and assigned an experiment attempting to emulate a real-life condition (Comparison of a fingerprint on a business card). The real-world accuracy was one unique element that this paper concentrated on, this was portrayed through the authors' attention to detail when designing the experiment.

Overall, results from the experiments emulated previously discussed literature. Textual representations faired effectively well with Words and Sentences again being some of the best options overall. Performance of graphical representation was mixed with OpenSSH having performance matching that of a strong textual mapping while Unicorns having the worst performance overall. This performance with graphical representations is consistent with the available literature apart from the positive performance of OpenSSH. The results of different modes of comparison were highly consistent with other studies where again the recommendation was to not use ``Compare-and-Select'' as its performance across the board was below even basic levels of security.

\textbf{M. Shirvanian \textit{et al.}}\cite{shirvanian2017pitfalls} produced further work in the context of secure messaging pairing. This paper was unique for a number of reasons. First was to consideration for ``remote-vs-proximity'' pairing where, this is the first consideration of this aspect found in the literature. Second, was their studies aim into investigating fingerprint security and usability in a end-to-end encryption context. The study compared Numerical, Visual and Auditory representations. This is a standard setup for similar studies in the literature. QR code verification was also added to the proximity tests. They split the variance of attack severity into three levels; one character change, one block change and changes to the entire fingerprint.

The findings from the paper showed a high false negative rate for all the schemes, this is an aspect also missing from the literature. Alongside this, results for usability were lower in a remote setting for all the schemes where the author comments on the expected nature of this result.\\
The paper measured the number of successful attacks in the False Accept Rate (FAR). With this in mind images were the most secure method of authentication in the remote setting, but voted as the method with the worst usability. This is highly inconsistent with previous work in this area. However, this could be due to the unique setting of remote verification resulting in distinct results from user studies. Without further work in this area it is difficult to conclusively validate these results. In the proximity setting the results were negligible where at most 2.67\% of attacks were successful. In summary, this paper provides a unique look into the real-world aspects of fingerprint verification in a end-to-end context providing a unique viewpoint on the usability and possible security of the main fingerprint encoding schemes.

In conclusion for this section, there has been extensive work in the investigation of the usability of textual and visual encoding schemes alongside methods of comparison. There are mixed opinions about visual representations, however, the majority defines them as easy to use but severely lacking in security. Research on textual representations, however, is in agreement; natural language based encodings such as words or sentences provide some of the highest usability and security over the baseline performance of alphanumerical encodings. The literature also concurs on the use of Compare-and-Confirm as the best method of comparing fingerprints overall, with Compare-and-Select being highlighted for its poor security and usability.\\
The starkest limitations of the literature are the range of participant and encoding entropy, where at its worst only 22-bits of entropy were tested. Another limitation is the scant amount of research solely investigating modes of comparison, where the literature only contains two papers with this aim. This is, therefore, an area that requires further research. A potential research avenue could be the analysis of comparison methods in unique scenarios, one example of this could be the idea of ``remote-vs-proximity'' proposed by M. Shirvanian \textit{et al.}\cite{shirvanian2017pitfalls}. Even with these limitations, as a whole the literature on this topic has provided a relatively comprehensive and coherent view on the usability of textual and visual hash encoding schemes with relatively clear recommendations on methods of comparison and suitable encoding schemes.

>>>>

\section{Introduction}
The increasing use of public key cryptography by instant messaging and secure email means key fingerprint 
verification is an ever more important task. One of the biggest risks to the security of the communication channel is a Man-in-the-middle (MiTM) attack . A successful MiTH attack can circumvent the encryption as it allows an attacker to read all the encrypted data. A countermeasure for this is the verification of each parties' fingerprint. \textbf{[...]}

% \textbf{TODO: } Finish introduction

% \textbf{TODO: } - Improve the flow of the first section, need to phrase it like I'm telling a story
% \textbf{TODO: } - Create written introduction setting the scene. \\
% \textbf{TODO: } - Add participant size to M. Shirvanian (25) \\
% \textbf{TODO: } - Create a background section for Compare-and-X
% \\
% \textbf{TODO: } - Create comparison area where inconsistencies are compared

\section{Literature Review}
\subsection{Fingerprint representation comparison}

Fingerprints are used to quickly compare the similarity of two keys. The comparison occurs between the digest of a strong one-way hash function. Historically these have been represented as a hexadecimal string and users were tasked with manually comparing fingerprints between two substrates, for example, a monitor screen and a business card. We will refer to this process throughout the paper as the \textit{``authentication ceremony''}. \\
It has been shown that the average human can only hold 7-digits worth of data in our working memory\cite{miller1956magical}. This rules out the possibility of comparing full digests. For examples SHA-1 is 40 hex digits (160-bit) and, therefore, will make it difficult for effective comparison. Therefore , if the  process is to involve humans interaction, there is a need for schemes that work effectively with our abilities and senses.

Research in this area has taken various encoding schemes and compared their fallibility to impersonated attacks. Impersonation of a key would involve brute force of the encoding schemes through the mass creation of keys. The main elements of comparison have been accuracy of attack detection and time to complete the comparison. In this context we will consider the metrics a measure of security and usability respectively.

\subsection*{Effective schemes}
Results 
from the literature consistently show the effectiveness of language based encodings (Words or Sentences). This was shown in \cite{dechand2016empirical}, \cite{tan2017can} and \cite{kainda2009usability} with the accuracy of these schemes all being above 94.00\%. In all cases these were the best scemes in terms of accuracy. The exception to this is the work performed by \textbf{Hsiao, \textit{et al.}}\cite{hsiao2009study} in 2009 with Words achieving an abnormal accuracy of 63.00\% (Comparisons of methodology will be discussed in later sections).
\\\\
Aside from textual representations were graphical schemes. Examples of schemes assessed were: Random Art\cite{perrig1999hash}, Flag\cite{ellison2003public}, T-Flag\cite{lin2010spate}, Vash\footnote{https://github.com/thevash/vash}, OpenSSH visual host key and Unicorns\footnote{https://unicornify.pictures/} to name a few. Accuracy from these schemes were mixed with positive results obtained by \textbf{Hsiao, \textit{et al.}}\cite{hsiao2009study} where their best accuracy was achived by Random Art (94.00\%) and Flag Ext.\footnote{The authors' own improvement on Flag}(88.00\%). On the other hand, the worst accuracy was obtained by Flag(50\%). The only other paper assessing graphical representations was the work of \textbf{Tan \textit{et al.}}\cite{tan2017can} where they also achieved mixed results with accuracies ranging from 46\% to 90\%.
\\
In terms of usability of graphical schemes the literature concurred on their ease-of-use. In terms of speed of comparison all schemes consistently obtained some of the quickest comparisons times. This is also supported indirectly from the the comparison results of alternative literature \cite{kainda2009usability}\cite{dechand2016empirical} where with these in consideration graphical schemes still remain the quickest overall.
In terms of research into the performance of graphical schemes, the literature does not contain an extensive review with only two papers including graphical schemes in their comparison. This is, therefore, a candidate for further research.

\subsection*{Experimentation comparison}
